Probability Aggregation: Dynamic Hierarchical Modeling of Expert Beliefs in Geopolitical Events

### 2013-08-26 paper.tex
Edits

\section{Discussion}
This paper began with an introduction of a rather unorthodox but nonetheless realistic time-series setting in which probability forecasts are made very infrequently by a heterogeneous group of experts. The resulting data is too sparse to be modeled efficiently with standard time-series methods.  In response to the lack of appropriate modeling procedures, our work introduces an interpretable time-series model that incorporates self-reported expertise to capture a sharp and well-calibrated estimate of the crowd belief. The model is estimated in two-steps: The first step, known as the \textit{sampling step}, samples constrained versions of the model parameters via a Gibbs sampler that only makes use of standard distributions (see Appendix \ref{appendix} for technical details). The second step, known as the \textit{calibration step}, involves a one-dimensional optimization procedure that transforms the constrained parameter values to their unconstrained counterparts.

\subsection{Summary of Findings}
The model was applied to an unusually large dataset on probability forecasting. The estimated crowd belief was found to be sharp and well-calibrated under both in- and out-of-sample probability estimation. This has direct implications  on predictive power and model interpretability. Firstly, the model was shown to outperform other probability aggregators in terms of forecasting ability. In particular, the model was deemed a well-balanced compromise that avoids over-fitting without begin overly conservative. Secondly, the crowd belief was used as the no-bias reference point to study the bias among groups of experts with different levels of self-reported expertise. All the groups were found to be under-confident. The under-confidence, however, decreased as the level of self-reported expertise increased. This result is about groups of experts and hence does not conflict with the well-known result of the individuals being over-confident (see, e.g.,  \citet{lichtenstein1977calibration, morgan1992uncertainty, bier2004implications}). Besides making predictions or studying group-level bias, the model can be used to generate estimates of many problem-specific quantities. These quantities have clear interpretations and can be combined in many interesting ways to answer a range of hypotheses about different types of questions and expert behavior.

\subsection{Limitations and Directions for Future Research}
This paper developed a time-series procedure that addresses the main challenges in modeling sparse probability forecasting data. The introduced model was kept parsimonious and can be viewed as a basic template that is amenable to many future extensions. For instance, most of the model parameters were assumed to remain constant over time. It is intuitively reasonable, however, that the behavior of the model parameters is different earlier as opposed to later on in the question.  For instance, the level of disagreement, $\sigma^2_k$, among the experts can be expected to decrease towards the final time point when the the question finally resolves. This hypothesis could be investigated by letting $\sigma^2_{t,k}$ depend on the previous term $\sigma^2_{t-1,k}$.

If personal information on the experts were inquired during the data collection process, it is possible to study the forecasting behavior of different kinds of experts. For instance, in this paper the multiplicative bias was modeled as a function of self-reported expertise. The model, however, is by no means restricted to the study of self-reported expertise. Parameter dependencies could be constructed based on many other expert characteristic, such as gender, education, or specialty, to produce a range of novel insight on the forecasting behavior across different groups of experts. Furthermore, expert characteristics could be interacted with question types, such as economic, domestic, or international, to learn about the kind of experts who generally perform well on certain types of questions. This would be of interest to the decision-maker who could use the information as a basis for consulting only a high-performing subset of the available experts.

>>>>

\section{Discussion}
This paper began with an introduction of a rather unorthodox but nonetheless realistic time-series setting where probability forecasts are made very infrequently by a heterogeneous group of experts. The resulting data is too sparse to be modeled well with standard time-series methods.  In response to this lack of appropriate modeling procedures, our work introduces an interpretable time-series model that incorporates self-reported expertise to capture a sharp and well-calibrated estimate of the crowd belief. The model estimation is performed in two steps. The first step, known as the \textit{sampling step}, samples constrained versions of the model parameters via a Gibbs sampler. The sampling is done from standard distributions with fast convergence to the target distribution (see Appendix \ref{appendix} for technical details). The second step, known as the \textit{calibration step}, uses a one-dimensional optimization procedure to transforms the constrained parameter values to their unconstrained counterparts. With minimal sacrifices in model interpretability and efficiency, this procedure is able to extend forecasting literature towards rather unexplored areas of probability aggregation.

\subsection{Summary of Findings}
The model was applied to an unusually large dataset on expert probability forecasts. The estimated crowd belief was found to be sharp and well-calibrated under both in- and out-of-sample settings. This has direct implications  on predictive power and model interpretability. Firstly, the model was shown to outperform other probability aggregators in terms of forecasting ability. In particular, the model was deemed a well-balanced compromise that avoids over-fitting without begin overly conservative. Secondly, the crowd belief was used as the no-bias reference point to study the bias among groups of experts with different levels of self-reported expertise. All the groups were found to be under-confident. The under-confidence, however, decreased as the level of self-reported expertise increased. This result is about groups of experts and hence does not conflict with the well-known result of the individuals being over-confident (see, e.g.,  \citet{lichtenstein1977calibration, morgan1992uncertainty, bier2004implications}). Besides making predictions or studying group-level bias, the model can be used to generate estimates of many problem-specific parameters. These quantities have clear interpretations and can be combined in many interesting ways to explore a range of hypotheses about different types of questions and expert behavior.

\subsection{Limitations and Directions for Future Research}
Our model pertains model parsimony while addressing the main challenges that arise from modeling sparse probability forecasting data. Therefore it can be viewed as a basis for many future extensions. To illustrate some potential directions, recall that most of the model parameters are assumed to remain constant over time. It is intuitively reasonable, however, that these model parameters behave differently during different time intervals of the question.  For instance, the level of disagreement (represented by $\sigma^2_k$ in our model) among the experts can be expected to decrease towards the final time point when the the question finally resolves. This hypothesis could be explored by letting $\sigma^2_{t,k}$ evolve dynamically as a function of the previous term $\sigma^2_{t-1,k}$ and random noise.

If personal information on the experts were inquired during the data collection process, it may be of interest to study the forecasting behavior of different kinds of experts. For instance, this paper modeled the bias separately within each expertise group. This, however, is by no means restricted to the study of bias or its relation to self-reported expertise. Different parameter dependencies could be constructed based on many other expert characteristic, such as gender, education, or specialty, to produce a range of novel insight on the forecasting behavior across different groups of experts. Furthermore, expert characteristics could be interacted with question types, such as economic, domestic, or international, to learn about the kind of experts who generally perform well on certain types of questions. This would be of interest to the decision-maker who could use the information as a basis for consulting only a high-performing subset of the available experts.
